---
title: "Coursera Statistical Inference Course Project: Part 2"
author: "Brad Tweardy"
date: "Saturday, March 21, 2015"
output: pdf_document
---

Introduction
================================================================================
In this project we will analyze the ToothGrowth data in the R datasets package.

Basic Analysis
================================================================================
Below we'll load the dataset and provide a basic analyis of it.
```{r, message = FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
```
```{r}
data(ToothGrowth)
ggplot(data = ToothGrowth, aes(x = dose, y = len, fill = supp))+
        geom_bar(stat = "identity") +
        facet_grid(.~supp) +
        xlab("Dosage of Supplement (in milligrams)") +
        ylab("Tooth Length") +
        ggtitle("ToothGrowth Plot by Supplement")
```
we can deduct from this data that a higher dosage of either supplement 
correlates to longer teeth.

Basic Summary
================================================================================
```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
summary(ToothGrowth)
table(ToothGrowth$supp, ToothGrowth$dose)
```
We see that there are three dosage amounts and two supplement types. We're also able to verify that the tests are distributed equally by supplement and dosage.

Comparison of Tooth Growth by Supplement and Dosage
================================================================================
```{r}
VC_t <- t.test(len~supp, paired = FALSE, var.equal = TRUE, data = ToothGrowth)
OJ_t <- t.test(len~supp, paired = FALSE, Var.equal = FALSE, data = ToothGrowth)
comp_supp <- data.frame("p-value" = c(VC_t$p.value, OJ_t$p.value), 
                        "Conf-Low"  = c(VC_t$conf[1], OJ_t$conf[1]),
                        "Conf-High" = c(VC_t$conf[2], OJ_t$conf[2]), 
                        row.names = c("Equal Variance","Unequal Variance"))
comp_supp
```
We see that the p value of Equal and Unequal variables are higher than five percent. We also see that Confidence Intervals in both t tests contain zero. With this data we cannot ignore the null hypothesis tests. 
```{r}
dose_.5 <- ToothGrowth[which(ToothGrowth$dose ==.5), 1]
dose_1.0 <- ToothGrowth[which(ToothGrowth$dose ==1), 1]
dose_2.0 <- ToothGrowth[which(ToothGrowth$dose ==2), 1]
Tcomp_.5_1.0 <- t.test(dose_.5, dose_1.0, paired = F, var.equal = T)
Fcomp_.5_1.0 <- t.test(dose_.5, dose_1.0, paired = F, var.equal = F)
Result_comp_.5_1.0 <- data.frame("p-value" = c(Tcomp_.5_1.0$p.value,
                                             Fcomp_.5_1.0$p.value), 
                                 "Conf-Low" = c(Tcomp_.5_1.0$conf[1], 
                                                Fcomp_.5_1.0$conf[1]),
                                 "Conf-High"=c(Tcomp_.5_1.0$conf[2],
                                               Fcomp_.5_1.0$conf[2]), 
                                 row.names=c("Equal Variance",
                                             "Unequal Variance"), 
                                 "Dose" = "0.5 to 1")
Tdose_.5_2.0 <- t.test(dose_.5, dose_2.0, paired = F, var.equal = T)
Fdose_.5_2.0 <- t.test(dose_.5, dose_2.0, paired = F, var.equal = F)
Result_comp_.5_2.0 <- data.frame("p-value" = c(Tdose_.5_2.0$p.value, 
                                               Fdose_.5_2.0$p.value),
                                 "Conf-Low"=c(Tdose_.5_2.0$conf[1],
                                              Fdose_.5_2.0$conf[1]), 
                                 "Conf-High"=c(Tdose_.5_2.0$conf[2],
                                               Fdose_.5_2.0$conf[2]), 
                                 row.names=c("Equal Var","Unequal Var"), 
                                 "Dose"="0.5 to 2")
Tdose_1.0_2.0 <- t.test(dose_1.0, dose_2.0, paired=F, var.equal=T)
Fdose_1.0_2.0 <- t.test(dose_1.0, dose_2.0, paired=F, var.equal=F)
Result_comp_1.0_2.0 <- data.frame("p-value" = c(Tdose_1.0_2.0$p.value, 
                                                Fdose_1.0_2.0$p.value),
                                  "Conf-Low"=c(Tdose_1.0_2.0$conf[1],
                                               Fdose_1.0_2.0$conf[1]),
                                  "Conf-High"=c(Tdose_1.0_2.0$conf[2],
                                                Fdose_1.0_2.0$conf[2]), 
                                  row.names=c("Equal Var","Unequal Var"), 
                                  "Dose"="1 to 2")
Result_Dose <- rbind(Result_comp_.5_1.0,Result_comp_.5_2.0,Result_comp_1.0_2.0)
Result_Dose
```
As we can see from the data above, we can confidently conclude that dosage levels have a impact on tooth growth. We can say that a higher dosage of either supplement will result in an increase in tooth growth. In supporting these findings, we can see in all performed tests that the p values are relatively low and the confidence intervals do not contain zeros.

Final Analysis
================================================================================
First, in order for this analysis to be accurate we must assume a few pieces of
information that was not given. In general, we have to assume that the overall health, diet, age and weight is relatively equal in all test subjects.

In conclusion, we can say for certain that the dosage amount effects the amount of tooth growth. Higher dosage normally results in higher length of tooth 
growth. What we cannot say is that one supplement over the other results in more
tooth growth.